Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 intro.tex
intro spell check

Current approaches to object class detection have reached a remarkable
level of performance in 2D bounding box
localization~\cite{pascal12, Felzenszwalb10, Overfeat, Krizhevsky12, Girshick14}, due to their
ability to generalize across differences in object appearance,
lighting, and viewpoint. While this generalization
ability of these methods is benefitial for robustness, it also limits
the level of detail that these detectors can deliver as an output.

@@

For (i), recent approaches to 2D-3D matching~\cite{Aubry14, Lim14}
typically rely on a
collection of 3D exemplar models, which they render from a
large number of viewpoints. The resulting artifial images are then
used to train exemplar models that can be matched to a real-world
image at test time.
%
For (ii), it has been realized that template-based exemplar detectors
based on HOG~\cite{Dalal05} features can be trained analytically, by
replacing the standard SVM with an LDA classifier
(E-LDA)~\cite{Hariharan12}. As a result, it is feasible to train
hundreds of thousands of mid-level patch detectors for
recognition~\cite{Aubry14}.
%
Unfortunately, the performance of E-LDA relies crucially on an
additional calibration step that equalizes the detection scores of
independently trained exemplar models. Since this step involves costly
mining of hard negative examples~\cite{Dalal05,Felzenszwalb10} on a
validation set, it constitutes the major computational bottleneck of
E-LDA, limiting its scalability.

In this paper, we propose a novel method for 2D-3D alignment of
exemplar CAD models to real-world images that circumvents the need for
calibration, and greatly enhances the scalability of E-LDA. As a
result, we can render novel views and train corresponding exemplar
models {\em on the fly}, without the need for offline processing. As a
by-product, we can formulate the alignment problem as a parameter
search in a continuous pose space, consisting of yaw, pitch and roll, resulting in truely continuous pose estimates, which
we implement using MCMC sampling.
%  require a large number of rendering and 2D
% templates for each of the CAD model with specific viewpoint. Since
% there are so many templates that it becomes almost impossible to get
% all the true positive image from training data. Instead \cite{Aubry14,
%   Lim14} used whitening as a preprocessing stage \cite{Hariharan12} to
% make templates more discriminative. Then calibration is done on a
% small training dataset.
%
% Although whitening to create template is faster than some of
% instance-level matching algorithms such as training Exemplar-SVM
% \cite{Malisiewicz11} or complicated models such as LEVAN
% \cite{Divvala14}, whitening and calibration large number of templates
% takes very long time. Also, once the fixed discrete viewpoint
% templates have been made and calibrated, it is difficult to fine-tune
% the prediction made using the large collection of templates.

@@

First, we present a novel method for training exemplar models from
rendered 3D CAD data {\em on the fly}, enabled by a novel variant of
E-LDA, and making efficient use of the specific characteristics of
rendered images. To our knowledge, our method constitutes the first
attempt to simultaneously render and train exemplar detectors on the
fly.
% In our paper, we present Non-Zero Whiten Histograms of Orientations
% (NZ-WHO) to dramatically speed up the template generation and get rid
% of time consuming calibration and yet achieves performance on par with
% calibrated templates. First, we will present NZ-WHO is more
% discriminative than WHO and introduce Conjugate Gradient method to
% speed up the whitening and create a high resolution template in just
% 80 milliseconds. Also, using this fast NZ-WHO template generation, we
% present \textbf{on-the-fly} template generation to fine tune an
% initial pose estimation to continuously refine translation, viewpoints
% (yaw, roll, pitch) and focal length.

>>>>

Current approaches to object class detection have reached a remarkable
level of performance in 2D bounding box
localization~\cite{pascal12, Felzenszwalb10, Overfeat, Krizhevsky12, Girshick14}, due to their
ability to generalize across differences in object appearance,
lighting, and viewpoint. While this generalization
ability of these methods is beneficial for robustness, it also limits
the level of detail that these detectors can deliver as an output.

@@

For (i), recent approaches to 2D-3D matching~\cite{Aubry14, Lim14}
typically rely on a
collection of 3D exemplar models, which they render from a
large number of viewpoints. The resulting artificial images are then
used to train exemplar models that can be matched to a real-world
image at test time.
%
For (ii), it has been realized that template-based exemplar detectors
based on HOG~\cite{Dalal05} features can be trained analytically, by
replacing the standard SVM with an LDA classifier
(E-LDA)~\cite{Hariharan12}, resulting in a whitened feature
representation, termed WHO. As a result, it is feasible to train
hundreds of thousands of mid-level patch detectors for
recognition~\cite{Aubry14}.
%
Unfortunately, the performance of WHO relies crucially on an
additional calibration step that equalizes the detection scores of
independently trained exemplar models. Since this step involves costly
mining of hard negative examples~\cite{Dalal05,Felzenszwalb10} on a
validation set, it constitutes the major computational bottleneck of
WHO, limiting its scalability.

In this paper, we propose a novel method for 2D-3D alignment of
exemplar CAD models to real-world images that circumvents the need for
calibration, and greatly enhances the scalability of WHO. As a
result, we can render novel views and train corresponding exemplar
models {\em on the fly}, without the need for offline processing. As a
by-product, we can formulate the alignment problem as a parameter
search in a continuous pose space, consisting of yaw, pitch and roll, resulting in truly continuous pose estimates, which
we implement using MCMC sampling.
%  require a large number of rendering and 2D
% templates for each of the CAD model with specific viewpoint. Since
% there are so many templates that it becomes almost impossible to get
% all the true positive image from training data. Instead \cite{Aubry14,
%   Lim14} used whitening as a preprocessing stage \cite{Hariharan12} to
% make templates more discriminative. Then calibration is done on a
% small training dataset.
%
% Although whitening to create template is faster than some of
% instance-level matching algorithms such as training Exemplar-SVM
% \cite{Malisiewicz11} or complicated models such as LEVAN
% \cite{Divvala14}, whitening and calibration large number of templates
% takes very long time. Also, once the fixed discrete viewpoint
% templates have been made and calibrated, it is difficult to fine-tune
% the prediction made using the large collection of templates.

@@

First, we present a novel method for training exemplar models from
rendered 3D CAD data {\em on the fly}, enabled by a novel variant of
WHO, termed NZ-WHO, and making efficient use of the specific
characteristics of rendered images. To our knowledge, our method
constitutes the first attempt to simultaneously render and train
exemplar detectors on the fly.
% In our paper, we present Non-Zero Whiten Histograms of Orientations
% (NZ-WHO) to dramatically speed up the template generation and get rid
% of time consuming calibration and yet achieves performance on par with
% calibrated templates. First, we will present NZ-WHO is more
% discriminative than WHO and introduce Conjugate Gradient method to
% speed up the whitening and create a high resolution template in just
% 80 milliseconds. Also, using this fast NZ-WHO template generation, we
% present \textbf{on-the-fly} template generation to fine tune an
% initial pose estimation to continuously refine translation, viewpoints
% (yaw, roll, pitch) and focal length.
